# Bacterial community analysis for surfaces in a university classroom
### **Meadow *et al.* "Bacterial communities on classroom surfaces vary with human contact"**

-----------------

This document contains all statistical analyses conducted for the manuscript. Note that due to the random iterative nature of some analyses (such as beta-rarefaction, DB-RDA & PCA) some of the figure parameters will change slightly during reanalysis, though core results will remain essentially unchanged. 

All data to reproduce analysis can be found here: `https://github.com/jfmeadow/Meadow_etal_Surfaces_Analysis`

-----------------

Load necessary ecological analysis libraries. 

```{r,cache=TRUE}
library(vegan)
library(labdsv)
```



Pull workspace from big OTU table created for Lillis Air data (Meadow et al. 2013, Indoor Air). Also removed a list of plants and 'no blast hits' as described in that manuscript. 

This beginning workspace contains: 

* `lillis.big.table`: OTU table created with bigger sequence dataset containing airborne bacteria. Swabs are a subset of those samples. 
* `swab.map`: metadata mapping file
* `big.streptos.names`: vector of plant chloroplast OTUs - these get removed.
* `big.nbh.names`: vector of OTUs with no GreenGenes hit to bacteria level - these are also excluded from analysis. 
* `lillis.big.taxa`: taxonomic assignments from GreenGenes database.


```{r,cache=TRUE}
load('lillis.RData')
source('functions.R')  #functions for formatting shortcuts.
```


Extract OTU table and remove plant sequences as well as sequences not identified as bacteria. 

```{r,cache=TRUE}
swab.table.tmp <- lillis.big.table[row.names(swab.map), 
  -which(colnames(lillis.big.table) %in% c(big.streptos.names, big.nbh.names))]
swab.taxa <- lillis.big.taxa[-which(row.names(lillis.big.taxa) %in% c(big.streptos.names, big.nbh.names)), ]
```

Some important metrics:

* total sequences: `r sum(swab.table.tmp)`
* total samples: `r nrow(swab.map)`
* total number of OTUs in this dataset: `r length(which(colSums(swab.table.tmp) > 0))`
* total number of OTUs in bigger dataset (including air): `r ncol(swab.table.tmp)`

Even the most depauperate has >4000 sequences, so rarefy to that level. 

```{r,cache=TRUE}
swab.table <- rrarefy(swab.table.tmp, 4000)
swab.table <- swab.table[, which(apply(swab.table, 2, sum) > 0)]
swab.taxa <- swab.taxa[colnames(swab.table), ]
```

After rarefaction, some important metrics: 

* total sequences: `r sum(swab.table)`
* total number of OTUs: `r ncol(swab.table)`

--------------------

### Results:

Are communities indicative of surface type? In other words, do we have 4 distinct communities on these 4 distinct surfaces?

```{r,cache=TRUE}
swab.can <- vegdist(swab.table, 'canberra')
adonis(swab.can ~ swab.map$type)
```

Yes. Definitely. 

Distance-based Redundancy Analysis (DB-RDA) for discriminant analysis as well as visualization: 

```{r,cache=TRUE}
swab.caps <- capscale(swab.can ~ swab.map$type)
anova(swab.caps)
```

This test is essentially identical, but with different iterations, so different p-value. Create a version of the DB-RDA that will have individual OTUs weighting samples:

```{r,cache=TRUE}
swab.caps2 <- capscale(swab.table ~ swab.map$type, distance='canberra')
anova(swab.caps2)
scaps2.plot <- plot(swab.caps2)
```


This is all to get a dataframe for each set of indicator taxa sets
the edit commands were used to manually change taxa names for plotting. 

```{r,cache=TRUE}
sp.caps <- data.frame(scaps2.plot$species[, c(1,2)])
chair.caps.all <- sp.caps[which(sp.caps$CAP1 > 0 & sp.caps$CAP2 > 0), ]
desk.caps.all <- sp.caps[which(sp.caps$CAP1 > 0 & sp.caps$CAP2 < 0), ]
wall.caps.all <- sp.caps[which(sp.caps$CAP1 < 0 & sp.caps$CAP2 < 0), ]
floor.caps.all <- sp.caps[which(sp.caps$CAP1 < 0 & sp.caps$CAP2 > 0), ]
```

Function to find extant OTUs

```{r,cache=TRUE}
hypot <- function(d.f) {
  sqrt.d.f <- sqrt((d.f[, 1]^2) + (d.f[, 2]^2))
  invisible(sqrt.d.f)
  }
```

Highlight those OTUs with strongest weights for each of the 4 groups. Started with top 10, then trimmed to significant indicators (from analysis below).

```{r,cache=TRUE}
hypot.chair <- hypot(chair.caps.all)
chair.tops <- chair.caps.all[rev(order(hypot.chair)), ][1:8, ]

hypot.desk <- hypot(desk.caps.all)
desk.tops <- desk.caps.all[rev(order(hypot.desk)), ][1:5, ]

hypot.wall <- hypot(wall.caps.all)
wall.tops <- wall.caps.all[rev(order(hypot.wall)), ][1:4, ]

hypot.floor <- hypot(floor.caps.all)
floor.tops <- floor.caps.all[rev(order(hypot.floor)), ][1:6, ]
```


Then save species scores for easier plotting. 

```{r,cache=TRUE}
sp.caps <- scores(swab.caps2)$species
```


Put taxon names on the OTUs. This requires the two functions intended to format taxonomy output. The output is a OTU classification dataframe with total abundance in the last column. 


```{r,cache=TRUE}
taxo <- cleanTaxo(makeTaxo(taxo.in = swab.taxa$taxa.names, otu.table = swab.table, split = ';'))
head(taxo)
```

Put names on the right points. 

```{r,cache=TRUE}
chair.tops$taxa <- taxo[row.names(chair.tops), 6]
desk.tops$taxa <- taxo[row.names(desk.tops), 6]
wall.tops$taxa <- taxo[row.names(wall.tops), 6]
floor.tops$taxa <- taxo[row.names(floor.tops), 6]

chair.tops$pos <- 1
desk.tops$pos <- 1
wall.tops$pos <- 1
floor.tops$pos <- 1

cap.txt <- data.frame(x=c(1.3, 1.1, -.8, -1), 
                      y=c(.75, -.2, -1, .75))
```


Check for significance of each indicator pointed out. This uses the `indval` function in the `labdsv` package - it essentially follows the procedure outlined in Dufrene & Legendre (1998). A few steps follow to format `indval` output for downstream analysis. 

```{r,cache=TRUE}
indic <- indval(swab.table, swab.map$type)

ch.ids <- paste('X',row.names(chair.tops), sep='')
de.ids <- paste('X',row.names(desk.tops), sep='')
wa.ids <- paste('X',row.names(wall.tops), sep='')
fl.ids <- paste('X',row.names(floor.tops), sep='')

chair.tops$indic <- indic$pval[ch.ids]
desk.tops$indic <- indic$pval[de.ids]
wall.tops$indic <- indic$pval[wa.ids]
floor.tops$indic <- indic$pval[fl.ids]

all.tops <- rbind(chair.tops, desk.tops, wall.tops, floor.tops)
all.tops$surface <- factor(c(rep('chair', nrow(chair.tops)), 
                      rep('desk', nrow(desk.tops)), 
                      rep('wall', nrow(wall.tops)), 
                      rep('floor', nrow(floor.tops))))
for(i in 1:nrow(all.tops)) {
  all.tops$chair.ra[i] <- 
		mean(swab.table[swab.map$type == 'chair', row.names(all.tops)[i]] / 
           sum(swab.table[swab.map$type == 'chair', ]))
	all.tops$desk.ra[i] <- 
		mean(swab.table[swab.map$type == 'desk', row.names(all.tops)[i]] / 
           sum(swab.table[swab.map$type == 'desk', ]))
	all.tops$wall.ra[i] <- 
		mean(swab.table[swab.map$type == 'wall', row.names(all.tops)[i]] / 
           sum(swab.table[swab.map$type == 'wall', ]))
	all.tops$floor.ra[i] <- 
		mean(swab.table[swab.map$type == 'floor', row.names(all.tops)[i]] / 
           sum(swab.table[swab.map$type == 'floor', ]))
		}

all.tops
```

Then write a table for manual documentation of individual OTUs. This goes into Table 1. 

```{r}
write.table(all.tops, file='surface_indicators.txt', sep='\t', quote=FALSE)
```

-------------

Put surface samples in context with gut, skin, soil, phyllosphere. First read in `sources.csv`; this is a Class-level OTU table from multiple source environments. The same approach was used in Kembel et al. (2012) to identify potential sources in airborne hospital bacterial assemblages. 


```{r,cache=TRUE}
## source environments
sources.tmp <- read.csv('sourceHabitatsBlastClass.csv')
sources <- sources.tmp[, c(2,4,3)]
names(sources) <- c('sample', 'species', 'abundance')

## prepare surface OTU table
swabClass <- aggregate(t(swab.table), by=list(taxo$class), FUN='sum')
row.names(swabClass) <- swabClass[, 1]
swabClass <- swabClass[, -1]
swabClass <- data.frame(t(swabClass/4000))
swabClass <- dematrify(swabClass)

## combine and make longform
sourcesAll.tmp <- rbind(swabClass, sources)
sourcesAll <- matrify(sourcesAll.tmp)

## make mapping factor
sourcesMap <- data.frame(env=rep('skin', nrow(sourcesAll)))
row.names(sourcesMap) <- rn <- row.names(sourcesAll)
sourcesMap$env <- as.character(sourcesMap$env)
sourcesMap$env[grep('Swab', rn)] <- 'classroom'
sourcesMap$env[grep('Fcsw', rn)] <- 'gut'
sourcesMap$env[grep('phyllosphere', rn)] <- 'phyllosphere'
sourcesMap$env[grep('soil', rn)] <- 'soil'
sourcesMap$env[grep('water', rn)] <- 'water'
sourcesMap$env <- as.factor(sourcesMap$env)

sourcesMap$bg <- 'transparent'
sourcesMap$bg[1:58] <- swab.map$bg


```

Then create ordination of these sources combined with surfaces samples. 

```{r,cache=TRUE}
sourcesPCA <- pca(sourcesAll)
plantskinPCA1 <- sourcesPCA$scores[, 1]
```

How much variance was explained by the axes?

```{r, cache=TRUE}
(sourcesPCA$sdev[c(1:2)]^2)/sum(sourcesPCA$sdev^2)
```


Plot results

```{r,cache=TRUE}
sourcesMap$col <- 1
for (i in 1:nlevels(sourcesMap$env)) {
  sourcesMap$col[sourcesMap$env == levels(sourcesMap$env)[i]] <- i
  }
plot(sourcesPCA, type='n')
points(sourcesPCA$scores[1:58, 1:2], pch=21, bg=sourcesMap$bg[1:58], cex=2)
points(sourcesPCA$scores[-c(1:58), 1:2], col=sourcesMap$col[-c(1:58)], pch=16)
legend('bottomleft', legend=c('Chair', 'Desk', 'Wall', 'Floor', levels(sourcesMap$env)[-1]), 
                               pch=c(rep(21, 4), rep(16, 5)), col=c(1,1,1,1,2:6), 
                               pt.bg=c('darkslateblue', 'goldenrod3', 'wheat', 'chocolate3'))
```

```{r,cache=TRUE}
boxType <- factor(swab.map$type, levels=c('chair', 'desk', 'wall', 'floor'))

library(IDPmisc)
par(mar=c(5,6,2,1))
boxplot(swab.map$plantskin ~ boxType, notch=TRUE,
	col=c('darkslateblue', 'goldenrod3', 'wheat', 'chocolate3'), 
	ylab='', las=1, pch=21, bg=c('wheat', 'goldenrod3'), pch=1.5)
par(xpd=TRUE)
Arrows(-.2, -.0078, -.2, -.4317, sh.col='gray20', sh.lwd=2.5) 
Arrows(-.2, -.4317, -.2, -.0078, sh.col='gray20', sh.lwd=2.5) 
mtext(c('more like Skin', 'more like Plants'), side=2, at=c(-.7, .3), line=3)
text(c(1,2,3,3.85), c(rep(.4, 4)), c('a', 'b', 'bc', 'c'), font=1, cex=1.4)

TukeyHSD(aov(swab.map$plantskin ~ swab.map$type))
```


This revised version of the figure contains information from the source PCA performed above. First some baggage from previous analysis. 

```{r}
library(IDPmisc)
boxType <- factor(swab.map$type, levels=c('chair', 'desk', 'wall', 'floor'))
boxTypeNum <- rep(1, nrow(swab.map))
boxTypeNum[boxType == 'chair'] <- 1
boxTypeNum[boxType == 'desk'] <- 2
boxTypeNum[boxType == 'wall'] <- 3
boxTypeNum[boxType == 'floor'] <- 4

chair.caps <- chair.tops
desk.caps <- desk.tops
wall.caps <- wall.tops
floor.caps <- floor.tops
```


```{r,cache=TRUE,fig.height=4,fig.width=10}
layout(matrix(c(1,2,3), 1, 3), widths=c(1,1,.75))
par(xpd=FALSE, mar=c(3,3,3,1), las=0)
plot(swab.caps, type='none', xaxt='n', yaxt='n', xlab='', ylab='')
mtext(c('CAP 1', 'CAP 2'), side=c(1,2), line=c(0.3, 0.1), adj=c(1,1))
points(swab.caps, 'sites', pch=21, bg=swab.map$bg, cex=2)
text(c(cap.txt$x), c(cap.txt$y), c('Chairs', 'Desks', 'Walls', 'Floors'), cex=1)
mtext('a', adj=0, font=2, cex=2)
plot(sp.caps, pch=16, cex=.4, col=8, xlim=c(range(sp.caps)), ylim=c(range(sp.caps)),
  xaxt='n', yaxt='n', xlab='', ylab='')
mtext(c('CAP 1', 'CAP 2'), side=c(1,2), line=c(0.3, 0.1), adj=c(1,1))
abline(h=0, v=0, lty=3, lwd=.7, col=1)
# rect(-1.5, -1.5, 2, 2)
par(xpd=TRUE)
points(chair.caps$CAP1, chair.caps$CAP2, pch=21, bg='darkslateblue', cex=1, lwd=.5)
points(desk.caps$CAP1, desk.caps$CAP2, pch=21, bg='goldenrod3', cex=1, lwd=.5)
points(wall.caps$CAP1, wall.caps$CAP2, pch=21, bg='wheat', cex=1, lwd=.5)
points(floor.caps$CAP1, floor.caps$CAP2, pch=21, bg='chocolate3', cex=1, lwd=.5)
text(chair.caps$CAP1, chair.caps$CAP2, as.character(chair.caps$taxa), cex=.7, pos=chair.caps$pos)
text(desk.caps$CAP1, desk.caps$CAP2, as.character(desk.caps$taxa), cex=.7, pos=desk.caps$pos)
text(wall.caps$CAP1, wall.caps$CAP2, as.character(wall.caps$taxa), cex=.7, pos=wall.caps$pos)
text(floor.caps$CAP1, floor.caps$CAP2, as.character(floor.caps$taxa), cex=.7, pos=floor.caps$pos)
mtext('b', adj=0, font=2, cex=2)
par(mar=c(4,5,3,1))
boxplot(swab.map$plantskin ~ boxType, notch=TRUE,
  col=c('darkslateblue', 'goldenrod3', 'wheat', 'chocolate3'), 
	ylab='', las=1, pch=21, cex=0, bg=c('wheat', 'goldenrod3'))
points(swab.map$plantskin ~ jitter(boxTypeNum, .7), 
	pch=21, bg=swab.map$bg)
par(xpd=TRUE)
Arrows(-.55, 0.2, -.55, -.7, sh.col='gray40', sh.lwd=1.5, width=2.5, size=.5) 
Arrows(-.55, -.7, -.55, 0.2, sh.col='gray40', sh.lwd=1.5, width=2.5, size=.5) 
mtext('Source PCA Score', side=2, at=-.25, line=3.1, col='gray30', cex=.8)
mtext(c('Skin', 'Plants'), side=2, at=c(-.85, .4), line=2.5)
text(c(1,2,3,3.85), c(rep(.4, 4)), c('a', 'b', 'bc', 'c'), font=1, cex=1.4)
mtext('c', adj=0, font=2, cex=2)
```

-----------------

Create subsets of data for Mantel tests. Ie, test for geographic location within the room. 

```{r,cache=TRUE}
st.c <- swab.table[swab.map$type == 'chair', ]
st.f <- swab.table[swab.map$type == 'floor', ]
st.d <- swab.table[swab.map$type == 'desk', ]
st.w <- swab.table[swab.map$type == 'wall', ]

sm.c <- swab.map[swab.map$type == 'chair', ]
sm.f <- swab.map[swab.map$type == 'floor', ]
sm.d <- swab.map[swab.map$type == 'desk', ]
sm.w <- swab.map[swab.map$type == 'wall', ]

mantel(vegdist(st.c, 'canberra'), vegdist(data.frame(sm.c$xcor, sm.c$ycor), 'euclid'))
mantel(vegdist(st.f, 'canberra'), vegdist(data.frame(sm.f$xcor, sm.f$ycor), 'euclid'))
mantel(vegdist(st.d, 'canberra'), vegdist(data.frame(sm.d$xcor, sm.d$ycor), 'euclid'))
mantel(vegdist(st.w, 'canberra'), vegdist(data.frame(sm.w$xcor, sm.w$ycor), 'euclid'))
mantel(vegdist(swab.table, 'canberra'), vegdist(data.frame(swab.map$xcor, swab.map$ycor), 'euclid'))

plot(vegdist(swab.table, 'canberra'), 
	vegdist(data.frame(swab.map$xcor, swab.map$ycor), 'euclid'))
cor.test(vegdist(swab.table, 'canberra'), 
	vegdist(data.frame(swab.map$xcor, swab.map$ycor), 'euclid'))
# Mantel statistic r: 0.01047 
      # Significance: 0.399 
```

This is never significant, which indicates that surface type makes more difference than location around the room (eg, proximity to door, window, ...).

And just to check, does it matter whether the walls were sampled high or low?

```{r,cache=TRUE}
sm.w$location2 <- factor(sm.w$location2)
adonis(vegdist(st.w, 'canberra') ~ sm.w$location2)
```

Nope. 

---------------



Make stacked barchart of OTU relative abundances. This takes the `taxo` table created above.

```{r,cache=TRUE}
t.phylum <- aggregate(t(swab.table), by=list(taxo$phylum), sum)
t.class <- aggregate(t(swab.table), by=list(taxo$class), sum)
t.order <- aggregate(t(swab.table), by=list(taxo$order), sum)
```

Then order everything for visual greatness.

```{r,cache=TRUE,fig.show='hide'}
#### phylum  1% cutoff
row.names(t.phylum) <- t.phylum$Group.1
t.phylum <- t.phylum[, -1]
t.phylum <- t.phylum[order(rowSums(t.phylum)), order(swab.map$type)]
nrows <- nrow(t.phylum)
nshow <- 7
nkeep <- nrows-nshow+1
t.phylum <- rbind(colSums(t.phylum[1:nkeep-1, ]), t.phylum[nkeep:nrows, ])
cols.8 <- c('#D73027', '#FC8D59', '#FEE090', '#FFFFBF', '#E0F3F8', '#91BFDB', '#9475B4', '#969A97')

# class  1% cutoff
row.names(t.class) <- t.class$Group.1
t.class <- t.class[, -1]
t.class <- t.class[order(rowSums(t.class)), order(swab.map$type)]
nrows <- nrow(t.class)
nshow <- 10
nkeep <- nrows-nshow+1
t.class <- rbind(colSums(t.class[1:nkeep-1, ]), t.class[nkeep:nrows, ])
t.class <- cbind(t.class[, 1:14][, rev(order(t.phylum[8, 1:14]))], 
				  t.class[, 15:29][, rev(order(t.phylum[8, 15:29]))], 
				  t.class[, 30:43][, rev(order(t.phylum[8, 30:43]))], 
				  t.class[, 44:58][, rev(order(t.phylum[8, 44:58]))] )
cols.11 <- c('#A50026', '#D73027', '#F46D43', '#FDAE61', '#FEE090', 
			'#E0F3F8', '#ABD9E9', '#74ADD1', '#4575B4', '#313695', '#969A97')

# order
row.names(t.order) <- t.order$Group.1
t.order <- t.order[, -1]
t.order <- t.order[order(rowSums(t.order)), order(swab.map$type)]
nrows <- nrow(t.order)
nshow <- 10
nkeep <- nrows-nshow+1
t.order <- rbind(colSums(t.order[1:nkeep-1, ]), t.order[nkeep:nrows, ])
t.order <- cbind(t.order[, 1:14][, rev(order(t.phylum[8, 1:14]))], 
				  t.order[, 15:29][, rev(order(t.phylum[8, 15:29]))], 
				  t.order[, 30:43][, rev(order(t.phylum[8, 30:43]))], 
				  t.order[, 44:58][, rev(order(t.phylum[8, 44:58]))] )
cols.11 <- c('#A50026', '#D73027', '#F46D43', '#FDAE61', '#FEE090', 
			'#E0F3F8', '#ABD9E9', '#74ADD1', '#4575B4', '#313695', '#969A97')
# create index for bar mid-points. 
bar <- barplot(as.matrix(t.order/4000), col=rev(cols.11))

####  this had to be run las to maintain phylum order in other charts
t.phylum <- cbind(t.phylum[, 1:14][, rev(order(t.phylum[8, 1:14]))], 
  			  t.phylum[, 15:29][, rev(order(t.phylum[8, 15:29]))], 
				  t.phylum[, 30:43][, rev(order(t.phylum[8, 30:43]))], 
				  t.phylum[, 44:58][, rev(order(t.phylum[8, 44:58]))] )


ph.names <- row.names(t.phylum)
ph.names[1] <- 'other (<1%)'
cl.names <- row.names(t.class)
cl.names[1] <- 'other (<1%)'
or.names <- row.names(t.order)
or.names[1] <- 'other (<4%)'
```

Figure 3. Taxonomy bar chart.

```{r,fig.height=8,fig.width=9}
layout(matrix(c(1,2,3, 7,1,2,3, 7,4,5,6, 7), 4, 3), heights=c(1,1,1,.1))
par(mar=c(1,4,2,1))
barplot(as.matrix(t.phylum/4000), col=rev(cols.8), border=rev(cols.8), xaxt='n', las=1)
mtext('Phylum', font=2, col='gray20')
segments(c(mean(bar[c(14,15)]), mean(bar[c(29,30)]), mean(bar[c(43,44)])), c(0,0,0),
	c(mean(bar[c(14,15)]), mean(bar[c(29,30)]), mean(bar[c(43,44)])), c(1,1,1))

mtext(c('chairs', 'desks', 'floors', 'walls'), side=1, 
	at=c(mean(bar[c(1,14)]), mean(bar[c(15,29)]), mean(bar[c(30,43)]), mean(bar[c(44,58)])), 
	font=2, col='gray20')  
barplot(as.matrix(t.class/4000), col=rev(cols.11), border=rev(cols.11), xaxt='n', las=1)
mtext('Class', font=2, col='gray20')
segments(c(mean(bar[c(14,15)]), mean(bar[c(29,30)]), mean(bar[c(43,44)])), c(0,0,0),
	c(mean(bar[c(14,15)]), mean(bar[c(29,30)]), mean(bar[c(43,44)])), c(1,1,1))
mtext(c('chairs', 'desks', 'floors', 'walls'), side=1, 
	at=c(mean(bar[c(1,14)]), mean(bar[c(15,29)]), mean(bar[c(30,43)]), mean(bar[c(44,58)])), 
	font=2, col='gray20')  
barplot(as.matrix(t.order/4000), col=rev(cols.11), border=rev(cols.11), xaxt='n', las=1)
mtext('Order', font=2, col='gray20')
segments(c(mean(bar[c(14,15)]), mean(bar[c(29,30)]), mean(bar[c(43,44)])), c(0,0,0),
	c(mean(bar[c(14,15)]), mean(bar[c(29,30)]), mean(bar[c(43,44)])), c(1,1,1))
par(xpd=TRUE)
mtext(c('chairs', 'desks', 'floors', 'walls'), side=1, 
	at=c(mean(bar[c(1,14)]), mean(bar[c(15,29)]), mean(bar[c(30,43)]), mean(bar[c(44,58)])), 
	font=2, col='gray20')  
par(mar=c(0,0,0,0))
plot(1,1, type='n', bty='n', xlab='', ylab='', xaxt='n', yaxt='n')
legend('left', legend=c(rev(ph.names)), pt.bg=cols.8, col=cols.8, bty='n', pch=22, cex=1.5, pt.cex=2.5)
plot(1,1, type='n', bty='n', xlab='', ylab='', xaxt='n', yaxt='n')
legend('left', legend=c(rev(cl.names)), pt.bg=cols.11, col=cols.11, bty='n', pch=22, cex=1.5, pt.cex=2.5)
plot(1,1, type='n', bty='n', xlab='', ylab='', xaxt='n', yaxt='n')
legend('left', legend=c(rev(or.names)), pt.bg=cols.11, col=cols.11, bty='n', pch=22, cex=1.5, pt.cex=2.5)
```















































